Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt

### 2024-10-16 Phase_II/draft/GAN.tex
..

\section{Generative Adversarial Networks}
Generative Adversarial Networks (GANs) were invented by Ian Goodfellow in 2014. The concept is rather simple, as it consists of two competing networks. The first network is called a generator and creates new images based on an input image. Here, for consistency, the images generated by the generator will be denoted as generated images. The second network is the discriminator, which tries to distinguish between the generated image and the real image (ground truth) \citep{goodfellow2014generative}.
As a consequence of training both networks alternately, the generated images eventually become indistinguishable from the real images. It is nothing more than a two-player min-max game, a famous problem in game theory. GAN was initially proposed in the formulation form as:
\begin{equation}
	\centering
	\begin{aligned}
		\min_{G} \max_{D} V(D, G) &= \mathbb{E}_{x \sim p_{data}(x)} \left[ \log D(x) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z)) \right) \right]
	\end{aligned}
	\label{eq:Basic_GAN}
\end{equation}
here, $V(D, G)$ is the value function of the min-max game.

@@

GAN can be extended to a conditional model \citep{mirza2014conditional}. In this case, both generator and discriminator receive extra information $y$, such that the value function of the conditional GAN (cGAN) can be expressed as:

@@

This type of network is shown to be very robust, as it has been applied to several problems. Examples include creating colored images from grayscale images, creating images of facades based on architectural labels, changing from day to night in different pictures, and predicting a map based on satellite data. A longer list of applications is given in \cite{isola2017image}.

>>>>

\section{Generative Adversarial Networks}
In 2014, Ian Goodfellow invented Generative Adversarial Networks (GANs). The concept is rather simple, as it consists of two competing networks. The first network is called a generator and creates new images based on an input image. Here, for consistency, the images generated by the generator will be denoted as generated images. The second network is the discriminator, which tries to distinguish between the generated image (predicted image) and the real image (ground truth) \citep{goodfellow2014generative}.
As a consequence of training both networks alternately, the generated images eventually become indistinguishable from the real images. It is nothing more than a two-player min-max game, a famous problem in game theory. GAN was initially proposed in the formulation form as:
\begin{equation}
	\centering
	\begin{aligned}
		\min_{G} \max_{D} V(D, G) &= \mathbb{E}_{x \sim p_{data}(x)} \left[ \log D(x) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z)) \right) \right]
	\end{aligned}
	\label{eq:Basic_GAN}
\end{equation}
here, $V(D, G)$ is the value function of the min-max game.

@@

Later, GAN extended its version to a conditional model \citep{mirza2014conditional}. In this case, both generator and discriminator receive extra information $y$, such that the value function of the conditional GAN (cGAN) can be expressed as:

@@

This type of network is shown to be very robust, as it has been applied to several problems. Examples include creating colored images from grayscale images based on architectural labels, changing from day to night in different pictures, and predicting a map based on satellite data. A longer list of applications is given in \cite{isola2017image}.

### 2024-10-16 Phase_II/draft/II.tex
..

\subsection{The signal for II}\label{sec:signal}
In a simple case, an II consists of two light collectors, for example, MAGIC-I and MAGIC-II, which are pointed to a star and measure the intensity $I_1(t)$ and $I_2(t)$ respectively \citep{acciari2020optical, dravins2013optical}. The signals from both detectors are cross-correlated and averaged over time, which results in an equation form
\begin{equation}
	g^{(2)} = \frac{\left\langle I_1(t) \cdot I_2(t + \tau) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} 
	\label{eqn:HBT}
\end{equation}
where $\tau$ is the time delay between both telescopes. If the light source is chaotic and randomly polarized, we have:
\begin{equation}
	g^{(2)} = 1 + \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eqn:HBT2}
\end{equation}
here, $\Delta f$ is the electronic bandwidth, $\Delta \nu$ is the optical bandwidth, and $\Delta f \ll \Delta \nu$ \citep{acciari2020optical}. $V_{12}$ is the Fourier transform of the source brightness distribution and is called complex visibility. Eqn.~\ref{eqn:HBT2} carry the information about the source's structure and shape. However, the phase information is lost due to the involvement of the absolute value of $V_{12}$. In observational astronomy, the correlation is often expressed in terms of the normalized contrast, given by:
\begin{equation}
	c = \frac{\left\langle \left( I_1(t) - \left\langle I_1 \right\rangle \right) \cdot \left( I_2(t + \tau) - \left\langle I_2 \right\rangle \right) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} = g^{(2)} - 1
\end{equation}
where, $\left\langle I_1 \right\rangle$ and $\left\langle I_2 \right\rangle$ are the mean intensity from two telescopes.

@@

The significance of the signal can be expressed as the signal-to-noise ratio (SNR), which depends on many factors. However, most importantly, it is proportional to the absolute value of visibility, which itself depends on the distance between the telescopes \citep{acciari2020optical} 
\begin{equation}
	SNR = A \cdot \alpha \cdot q \cdot n \cdot \abs{V_{12}}^{2}(d) \cdot \sqrt{b_v} \cdot F^{-1} \cdot \sqrt{\frac{T}{2}} \cdot \sigma.
	\label{eq:SNR}
\end{equation}
Here, $A$ is the total mirror area, $\alpha$ the quantum efficiency of the PMTs, $q$ the quantum efficiency of the optics, $n$ the differential photon flux from the source, and $b_v$ the cross-correlation bandwidth. The noise of the PMTs is accounted with $F$, $T$ denotes the observation time, and $\sigma$ is the normalized spectral distribution of the light (including filters) \citep{acciari2020optical}. While most of the parameters can be optimized with hardware, the only way to obtain a better SNR is to increase the observation time with a fixed number of telescopes.  
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/baseline.png}
	\caption{The tracking of baselines with four telescopes arranged in fig.~\ref{fig:teles} for one night of observation.}
	\label{fig:base}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{fig/ellipse/ellipse6018.jpg}
	\caption{This figure shows the simulated fast-rotating star. The brightness is highest at the pole, and gravitational darkening is present at the equator.}
	\label{fig:image}
\end{figure}
\subsection{Baseline considerations}
The measurement of the size of stellar objects through absolute visibility depends on the distance between the telescopes, which is called the baseline $B$. 
\begin{equation}
	V_{12}(B) = \frac{c(B)}{c(0)}
	\label{eq:angular_size_meas}
\end{equation}
However, this work needs a good SNR value for high precision of measurement, so the large covered observational plane with telescopes is a necessity for II \citep{acciari2020optical, abeysekara2020demonstration}. If the source is at the zenith, the coordinates in the Fourier plane ($u,v$) are given by:
\begin{equation}
	(u,v) = \frac{1}{\lambda} (B_N, B_E)
\end{equation}
where $B_N$ and $B_E$ are the baselines expressed in north and east coordinates. Since not all sources are at the zenith, and the telescopes are stationary, Earth's rotation plays an important role in covering the maximum observational plane through the rotated baselines, which is given as  
\begin{equation}
	\begin{pmatrix} u\\ v\\ w\\ \end{pmatrix} = R_x(\delta) \cdot R_y(h) \cdot R_x(-l) \begin{pmatrix} B_N\\ B_E\\ B_A\\ \end{pmatrix}
	\label{eq:baseline_rot}
\end{equation}
It traces an ellipse for every pair of telescopes. Furthermore, the different altitudes of the telescopes $B_A$ must be considered \citep{dravins2013optical, saha2020theory}. Here, $\delta$ is the declination, and $h$ is the hour angle of the stellar source, and $l$ is the latitude of the telescopes. The three matrices $R_i$ correspond to the fundamental representation of the SO(3) group \citep{saha2020theory}. Fig.~\ref{fig:base} shows the track of six baselines generated from the telescopes (fig.~\ref{fig:teles}) due to Earth's rotation. Since every pair of telescopes traces an ellipse in the Fourier plane, the total number of ellipses scales as follows:
\begin{equation}
	\label{eq:N_telescopes}
	\mathcal{N} = \frac{N_T \cdot (N_T -1)}{2}
\end{equation}

@@

Intensity Interferometry (II) detects the photon from the stellar objects, and the correlation of it results in squared visibility (explained in subsection.\ref{sec:signal}). According to Van Cittert Zernike's theorem, this signal is the Fourier transform of brightness distribution in the sky. Fig.~\ref {fig:ft} visualize this statement for the source shown in Fig.~\ref{fig:image} using II on linear and logarithmic scales. However, current observational techniques are unable to capture the full signal, and only partial data can be collected with a single night's simulation, which has been shown in Fig.~\ref{fig:ft_base}. It is the covered observational plane using the baselines shown in Fig.~\ref{fig:teles} on linear and logarithmic scales, respectively. However, the upcoming CTAO can promise to greatly expand the observable portion of the visibility plane for stellar objects, offering new insights into the mysteries of the universe.

>>>>

\subsection{The signal for II}\label{sec:signal}
In a simple case, an II consists of two light collectors, for example, MAGIC-I and MAGIC-II, which are pointed to a star and measure the intensity $I_1(t)$ and $I_2(t)$ respectively \citep{acciari2020optical, dravins2013optical}. The signals from both detectors are cross-correlated and averaged over time, which results in an equation form
\begin{equation}
	g^{(2)} = \frac{\left\langle I_1(t) \cdot I_2(t + \tau) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} 
	\label{eqn:HBT}
\end{equation}
where $\tau$ is the time delay between both telescopes. If the light source is chaotic and randomly polarized, we have:
\begin{equation}
	g^{(2)} = 1 + \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eqn:HBT2}
\end{equation}
here, $\Delta f$ is the electronic bandwidth, $\Delta \nu$ is the optical bandwidth, and $\Delta f \ll \Delta \nu$ \citep{acciari2020optical}. $V_{12}$ is the Fourier transform of the source brightness distribution and is called complex visibility. Eqn.~\ref{eqn:HBT2} carry the information about the source's structure and shape. However, it loses the phase information due to the presence of only the absolute value of $V_{12}$. In observational astronomy, the correlation in intensity is often expressed in terms of the normalized contrast, given by :
\begin{equation}
	c = \frac{\left\langle \left( I_1(t) - \left\langle I_1 \right\rangle \right) \cdot \left( I_2(t + \tau) - \left\langle I_2 \right\rangle \right) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} = g^{(2)} - 1
\end{equation}
where, $\left\langle I_1 \right\rangle$ and $\left\langle I_2 \right\rangle$ are the mean intensity from two telescopes.

@@

The significance of the signal is expressed as the signal-to-noise ratio (SNR), which depends on many factors. However, most importantly, SNR is proportional to the absolute value of visibility, which itself depends on the distance between the telescopes \citep{acciari2020optical} 
\begin{equation}
	SNR = A \cdot \alpha \cdot q \cdot n \cdot \abs{V_{12}}^{2}(d) \cdot \sqrt{b_v} \cdot F^{-1} \cdot \sqrt{\frac{T}{2}} \cdot \sigma.
	\label{eq:SNR}
\end{equation}
Here, $A$ is the total mirror area, $\alpha$ the quantum efficiency of the PMTs, $q$ the quantum efficiency of the optics, $n$ the differential photon flux from the source, and $b_v$ the cross-correlation bandwidth. The noise of the PMTs is accounted with $F$, $T$ denotes the observation time, and $\sigma$ is the normalized spectral distribution of the light (including filters) \citep{acciari2020optical}. While most of the parameters can be optimized with hardware, the only way to obtain a better SNR is to increase the observation time with a fixed number of telescopes.  
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/baseline.png}
	\caption{The tracking of baselines with four telescopes arranged in fig.~\ref{fig:teles} for one night of observation.}
	\label{fig:base}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{fig/ellipse/ellipse6018.jpg}
	\caption{This figure shows one of the simulated fast-rotating stars. The brightness is highest at the pole and lowest at the equator due to the presence of high-speed rotation of the star. This effect is named gravitational darkening in astronomy.}
	\label{fig:image}
\end{figure}
\subsection{Baseline considerations}
The measurement of the size of stellar objects through absolute visibility depends on the distance between the telescopes, which is called the baseline $B$
\begin{equation}
	V_{12}(B) = \frac{c(B)}{c(0)}.
	\label{eq:angular_size_meas}
\end{equation}
However, this work needs a good SNR value for high precision of measurement, so the large covered observational plane with telescopes is a necessity for II \citep{acciari2020optical, abeysekara2020demonstration}. If the source is at the zenith, the coordinates in the Fourier plane ($u,v$) are given by:
\begin{equation}
	(u,v) = \frac{1}{\lambda} (B_N, B_E)
\end{equation}
where $B_N$ and $B_E$ are the baselines expressed in north and east coordinates. Since not all sources are at the zenith, and the telescopes are stationary, Earth's rotation plays an important role in covering the maximum observational plane through the rotated baselines, which is given as  
\begin{equation}
	\begin{pmatrix} u\\ v\\ w\\ \end{pmatrix} = R_x(\delta) \cdot R_y(h) \cdot R_x(-l) \begin{pmatrix} B_N\\ B_E\\ B_A\\ \end{pmatrix}
	\label{eq:baseline_rot}
\end{equation}
It traces an ellipse for every pair of telescopes. Furthermore, the different altitudes of the telescopes $B_A$ must be considered \citep{dravins2013optical, saha2020theory}. Here, $\delta$ is the declination, and $h$ is the hour angle of the stellar source, and $l$ is the latitude of the telescopes. The three matrices $R_i$ correspond to the fundamental representation of the SO(3) group \citep{saha2020theory}. Fig.~\ref{fig:base} shows the track of six baselines generated from the telescopes (fig.~\ref{fig:teles}) due to Earth's rotation. Since every pair of telescopes traces an ellipse in the Fourier plane, the total number of ellipses scales as follows:
\begin{equation}
	\label{eq:N_telescopes}
	\mathcal{N} = \frac{N_T \cdot (N_T -1)}{2}
\end{equation}

@@

Intensity Interferometry (II) detects the photon from the stellar objects, and the correlation of it results in squared visibility (explained in subsection.\ref{sec:signal}). According to Van Cittert Zernike's theorem, this signal is the Fourier transform of brightness distribution in the sky. Fig.~\ref {fig:ft} visualize this statement for the source shown in Fig.~\ref{fig:image} using II on linear and logarithmic scales. However, current observational techniques are unable to capture the full signal, and only partial data can be collected with a single night's simulation, which has been shown in Fig.~\ref{fig:ft_base}. It is the covered observational plane using the baselines shown in Fig.~\ref{fig:teles} on linear and logarithmic scales, respectively. However, the instrumentation of existing and upcoming CTAO shows the possibility of covering the maximum observable portion of the visibility plane for stellar objects, offering high resolution to understand the stellar structure and its evolution with time.

### 2024-10-16 Phase_II/draft/ParaGan.tex
..

\section{The Parameter Analysis for GAN}
\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GAN, where the generator and discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks converge properly.

>>>>

\section{The Parameter Analysis for GAN}
\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GAN, where the generator and discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks are in good term to train the model.

### 2024-10-16 Phase_II/draft/Result.tex
..

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results, accurately providing visual information about the source's size and shape, as well as the distributed brightness. This image reconstruction was achieved using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover more (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate in collaboration with existing CTA. 
\subsection{Evaluation of GAN using moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity of the source \citep{teague1980image}. These parameters fully describe the ellipse that fits the image data based on the moments.

>>>>

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first, middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results. It accurately provides visual information about the source's size, shape, and brightness distribution over the surface. It is achieved here using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover maximum (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate. 
\subsection{Evaluation of GAN using moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help to quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image only. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity with respect to the center of the source \citep{teague1980image}. These parameters based on the moments fully describe the two-dimensional ellipse that fits the image data.

### 2024-10-16 Phase_II/draft/conclusion.tex
..

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry on optical wavelength. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes, equipped with large aperture areas and photon detectors capable of resolving signals within a nanosecond, are instrumental in capturing high-resolution images \citep{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now also being used for II, observing stellar objects with its 17-meter diameter mirror \citep{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \citep{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \citep{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts, such as the Southern Cherenkov Telescope Array (CTA) layout, to evaluate the impact on image reconstruction quality.

While the results of this study demonstrate the significant potential of machine learning, particularly cGAN, for image reconstruction in II, several aspects need further refinement. Detector efficiencies, which impact the signal-to-noise ratio (SNR) in real-world scenarios, have yet to be incorporated, and addressing them will be crucial for more accurate SNR estimation. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR and the robustness of the GAN in real-world applications. While cGAN has proven effective in this study, there are other methods for generating images from image data. Exploring and comparing alternative methods, as well as adapting the GAN architecture itself, could offer improvements. For instance, different loss functions could be implemented and compared to enhance reconstruction quality. Further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

>>>>

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry with optical wavelength range. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes are instrumental in capturing high-resolution images equipped with large aperture areas and sensitive photon detectors (capable of resolving signals within a nanosecond) \citep{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now also being used for II, observing stellar objects with its 17-meter diameter mirror \citep{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \citep{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \citep{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts to study the impact on image reconstruction quality, such as the Southern Cherenkov Telescope Array (CTA) with MAGIC.

While the results of this study demonstrate the significant potential of machine learning, particularly cGAN, for image reconstruction in II, several aspects need further refinement. (1) Detector efficiencies, which impact the signal-to-noise ratio (SNR) of real observational data, have yet to be incorporated. It will be crucial to address these challenges for more accurate SNR estimation. Incorporating these efficiencies in future work will be important for providing a more accurate estimate of the SNR. (2) Exploring and comparing different methods to generate images, which works better than cGAN to reconstruct stellar images with II. (3) Different loss functions can be used to check the reconstruction quality. So, there are further testing is required to refine the GAN and make it more robust and reliable. However, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

### 2024-10-16 Phase_II/draft/intro.tex
..

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in the N-body system. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. Yet, despite these technological advancements, II still faces a significant limitation: the absence of phase information, which is essential for the complete reconstruction of the brightness distribution of stellar sources.

@@

Currently, artificial intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using GAN \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. With the introduction of more telescopes, we can expect even greater precision, especially in recovering higher-order image moments. This approach holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.

>>>>

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in single/binary star systems. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique in comparison to the limitation of amplitude interferometry for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. However, despite these technological advancements, II faces a significant limitation in terms of phase loss and the absence of complete reconstruction of the brightness distribution of stellar sources.

@@

Currently, artificial intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and provide high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using GAN \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. We expect even greater precision of image reconstruction by including more telescopes, especially to recover the higher-order image moments. It holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.
